---
title: "STA 490 Final Team Project: Data Source and Team Assignment"
author: "Cheng Peng"
date: "West Chester University"
output: 
  html_document: 
    toc: yes
    toc_depth: 4
    toc_float: yes
    code_folding: hide
    code_download: yes
    smooth_scroll: yes
    number_sections: yes
    theme: readable
---

<style type="text/css">
div#TOC li {
    list-style:none;
    background-image:none;
    background-repeat:none;
    background-position:0;
}
h1.title {
  font-size: 24px;
  color: DarkRed;
  text-align: center;
}
h4.author { 
  /* Header 4 - and the author and data headers use this too  */
  font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkRed;
  text-align: center;
}
h4.date { 
  /* Header 4 - and the author and data headers use this too  */
  font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkBlue;
  text-align: center;
}
h1 { 
    /* Header 3 - and the author and data headers use this too  */
    font-size: 22px;
    font-family: "Times New Roman", Times, serif;
    color: darkred;
    text-align: center;
}
h2 { 
    /* Header 3 - and the author and data headers use this too  */
    font-size: 18px;
    font-family: "Times New Roman", Times, serif;
    color: navy;
    text-align: left;
}
h3 { 
    /* Header 3 - and the author and data headers use this too  */
    font-size: 15px;
    font-family: "Times New Roman", Times, serif;
    color: darkred;
    font-face: bold;
    text-align: left;
}
h4 { 
    /* Header 4 - and the author and data headers use this too  */
    font-size: 18px;
    font-family: "Times New Roman", Times, serif;
    color: darkred;
    text-align: left;
}
   /* center maps using chunk option: fig.align='center' */
.html-widget {
    margin: auto;
}
</style>

```{r setup, include=FALSE}
if (!require("ltm")) {
   install.packages("ltm")
   library(ltm)
}
if (!require("psych")) {
   install.packages("psych")
   library(psych)
}
if (!require("cocron")) {
   install.packages("cocron")
   library(cocron)
}
if (!require("knitr")) {
   install.packages("knitr")
   library(knitr)
}
if (!require("car")) {
   install.packages("car")
   library(car)
}
if (!require("corrplot")) {
   install.packages("corrplot")
   library(corrplot)
}
knitr::opts_chunk$set(echo = TRUE,       
                      warning = FALSE,   
                      result = TRUE,   
                      message = FALSE,
                      comment = NA)
```

\

# Introduction

This is essentially a capstone team project which involves multi-item survey data, survey instrument validation, information aggregation of multi-item survey data, SAP drafting, EDA and regression modeling, statistical analysis reporting, and HTML presentation.


# Data Scources

Three data sets are suggested for the team projects.

## Employee Turnover Data

The data set and its description can be found at <https://pengdsci.github.io/datasets/#employeeturnover>.

The following five variables measure the personality. We can aggregate these variables to make a personality index using principal component analysis (PCA) method.

* **extraversion**  - (outgoing/energetic vs. solitary/reserved)
* **independence**  - Agreeableness (friendly/compassionate vs. critical/rational)
* **selfcontrol** - Conscientiousness (efficient/organized vs. extravagant/careless)
* **anxiety**  - Neuroticism (sensitive/nervous vs. resilient/confident)
* **novator** - openness to experience (inventive/curious vs. consistent/cautious)

```{r}
EmployeeTurnOver = read.csv("https://pengdsci.github.io/datasets/EmployeeTurnOver/turnover-data-set.csv")
#personalityItems = EmployeeTurnOver[,c("extraversion","independ","selfcontrol","anxiety","novator")]
#calculate Cronbach's Alpha
#cronbach.alpha(personalityItems)      
#res.pca <- prcomp(personalityItems, scale = TRUE)
#res.pca
#summary(res.pca)
```
\

## Employee Attrition Data

The data set and its description can be found at <https://pengdsci.github.io/datasets/#employeeattrition>.

The following five variables measure employee satisfaction from different perspectives.


* **EnvironmentSatisfaction**
* **JobSatisfaction**
* **RelationshipSatisfaction**
* **WorkLifeBalance**

```{r}
EmployeeAttrition = read.csv("https://pengdsci.github.io/datasets/EmployeeAttrition/EmployeeAttritionData.csv")
#OverallSatisfaction = EmployeeAttrition[, c("EnvironmentSatisfaction", "JobSatisfaction", "RelationshipSatisfaction", "WorkLifeBalance")]
#cronbach.alpha(OverallSatisfaction)
#attr.pca <- prcomp(OverallSatisfaction, scale = TRUE)
#attr.pca
#summary(attr.pca)
#cor(OverallSatisfaction)
```

\


# Project Ideas

The above two data sets are related to employee retention. The following are the potential questions you may consider when analyze these data sets.


## Employee Attrition Data

Practical questions related to the **Employee Attrition Data**.


**General Questions for EDA**

* Any missing value issues? Need imputations? justification?

* Is there any potential issue of multi-collinearity? 

* Is it possible to aggregate the information in these satisfaction related variables to create a new *combined satisfaction variable*.

* Explore inter-relationship between satisfaction variables.


**Primary Questions**

Identify intervenable factors that are associated with 

* employee attrition, 
* engagement, and 
* performance.


**Secondary Questions**

Identify potential intervenable factors that impact the level of employees' satisfaction in different perspectives.

* Job satisfaction
* Environmental satisfaction
* Relationship satisfaction
* Work-life balance satisfaction


**Other Findings**

Document all practically important findings that were not asked to address any of clients' initial questions and make relevant recommendations.



## Employee Turnover Data














